---
title: "Question_3"
author: "Huimin Lu"
date: "2022/1/12"
output:
  word_document: default
  pdf_document: default
---
```{r}
setwd("H:\\Lund Data analysing\\Exam2021")
```

# 3a
```{r}
library(car)
library(DescTools)
library(DHARMa)
library(ggplot2)
library(ggmosaic)
library(lattice)
library(lmtest)
library(MASS)
library(MuMIn)

```
I choose chi square test, because in this question both dependent variable and independent variable are one- categery variable. so chi-squre test is suitable.
```{r}
moth <- read.csv("moth.csv")
str(moth)
attach(moth)
moth<-table(hab,subs)
moth
a<-matrix(moth,nrow=2,ncol=2)
colnames(a) <- c("subs0", "subs1")
rownames(a)<-c("cold","warm")
moth<-a
moth
```

```{r}
moth.dat <- data.frame(subs=c("subs0","subs0","subs1","subs1"), hab=c("cold","warm","cold","warm"), N=c(moth[,1], moth[,2]))
# moth.dat
ggplot(moth.dat, aes(x=subs, y=N, fill=hab)) +
	geom_bar(stat="identity", position="dodge") +
	theme_bw()
```

```{r}
## 2. Assumptions of frequency tests.
#a,Input data must be counts.
#b.Categories must be mutually exclusive.
#Obviously, the data is counted, and a moth can only belong to a single sub-species, and the habitat the moth choose is either "warm" or "cold". 
#c.Expected counts per category must be > 5.
fdata<-moth
fexp<-matrix(NA,nrow(fdata),ncol(fdata))
for (i in 1:nrow(fdata)) {
	for (j in 1:ncol(fdata)) {
		fexp[i,j]<-(((sum(fdata[i,]))/sum(fdata))*((sum(fdata[,j]))/sum(fdata))*(sum(fdata)))
		}
	}
fexp# the result shows that expected count of each cell greater than 5.
#All assumptions fullfild.
```


```{r}
## 3. Chi-square test.
# moth
chisq.test(moth)
#The result is that p-value = 0.06241, which acccept the null hypothesis, that the sub-species and habitats the moths choose are independent. So occurrence of the two subspecies doesn't differs between habitats.
```

```{r}
## 4, Fisher's exact test
fisher.test(moth)
#p-value = 0.06043, means that subspecies and their choice of habitat is independent.
```

```{r}
## 5. G-test
GTest(moth,correct = "none")
#p-value = 0.02287 means subspecies and their choice of habitat is dependent.
```
Among these three tests, because the expected counts are very low, and the data is a 2*2 contingency table, Fisher-test is the best test to analyse this data. And the result of Fisher-test is that subspecies and their choice of habitat is independent.    
In conclusion, the occurrence of the two subspecies doesn't differ between habitats.

#3C
Illustration

```{r}
# moth.dat
ggplot(moth.dat, aes(x=subs, y=N, fill=hab)) +
 geom_bar(stat="identity", position="dodge") +
 xlab("Sub-species of moth") +
 ylab("Counts") +
 scale_x_discrete(labels=c("Sub-pecies0", "Sub-species1")) +
 scale_fill_manual(name="Habitat", labels=c(cold="Cold", warm="Warm"), 
 values=c(cold="#CCFF99", warm="#FFFF99")) +
 theme_bw()

```

#3b
```{r}
moth<- read.csv("moth.csv")
moth$hab <- factor(moth$hab)
str(moth)
```

```{r}
## 1,Explore data
#Relationship between sub-species and host plant density
plot(subs~host, data=moth)
abline(lm(subs~host, data=moth))
# In this plot, the distribution of the density of plant of the sites that moths choose is different between two sub-species.
# And the regression line has a trend of increase.
#
#Relationship between sub-species and climate of habitat
counts.hab <- data.frame(table(moth$hab, moth$subs))
ggplot(counts.hab) +
    geom_mosaic(aes(weight=Freq, x = product(Var1), fill = Var2)) +
    theme_bw()
#The proportion of "warm" and "cold" is similar. 
#But "sub0" is likely to choose cold habitat, and "sub1" is likely to choose warm habitat.

```


```{r}
## 2,Construct full model
full.model <- glm(subs~host*hab, data=moth, family=binomial)
```

```{r}
## 3,Model selection
options(na.action = "na.fail")
output <- dredge(full.model, rank=AIC)
output
# The best fitted model is "subs~host", and the second one is "subs~host+hab"

```
## 4,Construct best model
``` {r}
fit.moth <- glm(subs~host, data=moth, family=binomial)
fit.moth2 <- glm(subs~host+hab, data=moth, family=binomial)
#Second-ranked model
Anova(fit.moth2, type=2)
#There is a variable in second-ranked model not significant. 
#So we only choose the first-ranked model,"subs~host". 
```
## 5,
```{r}
#Independent data
dwtest(fit.moth)
#The result is not significant, which means there is no correlation among data.

#Residuals do not deviate from predicted distribution
par(mfrow=c(2,2))
plot(fit.moth)

par(mfrow=c(1,1))
simulationOutput <- simulateResiduals(fittedModel=fit.moth, n=250)
plot(simulationOutput)

testDispersion(simulationOutput)

```



```{r}
## 6,Get results of binomial GLM
#First-rankd model
Anova(fit.moth, type=2)
#The difference of host plant density between two types is significant.

summary(fit.moth)
#the model is not significant, but the p-value is close to 0.05.
```
In conclusion, using GLM, the system compares of all possible model, fount that the model "subs~host" is the best fitted, which means that habitats couldn't determine the occurrence of two subspecies. Though the model "subs~host" is not significant, the plot shows that the occurrence of the sub-species is determined by the availability of the potential host plant. 


#3c
```{r}
## Illustration
plot(subs~host, xlab="Density of host plant", ylab="Sub-species", ylim=c(0,1.4), data = moth)
#
coefs<-coefficients(fit.moth)
coefs
x <- seq(min(moth$host), max(moth$host), len=100)
y <- coefs[1] + coefs[2]*x
pred <- exp(y)/(1+exp(y)) 
#
lines(x, pred, col="#339966", lty=2)

```





