---
title: "Q4"
author: "Huimin Lu"
date: "2022/1/15"
output:
  word_document: default
  pdf_document: default
---

```{r}
setwd("H:\\Lund Data analysing\\Exam2021")
```

```{r}
library(car)
library(GGally)
library(ggplot2)
library(gridExtra)
library(Hmisc)
library(lmtest)
library(MVN)
```

# 4a
The requirement of this question is that obtain a new variable to measure overall size. PCA allows to see the overall "shape" of the data. The score that gets through running PCA is that variable taht satisfied the requirement.
```{r}
#PCA
#Import data and check
morphs <- read.csv("morphs.csv")
morphs$morph <- factor(morphs$morph)
str(morphs)
```

```{r}
#morphs
#Check assumptions
#Sufficient sample size
str(morphs)
#68 samples, greater than 5*5=25.

#Correlations among variables
rcorr(as.matrix(morphs[,2:6]))

#Linearity
pairs(morphs[,2:6])
#Using ggplot2
ggpairs(morphs[,2:6])
#in the plot, the data fullfil the linearty

#No outliers
out <- NULL
for (i in 2:6) {
	out1 <- morphs[,i]>=mean(morphs[,i])+3*sd(morphs[,i]) | morphs[,i]<=mean(morphs[,i])-3*sd(morphs[,i])
	out2 <- which(out1==TRUE)
	out <- c(out, out2)
	out <- unique(out)
}
out
#Remove outliers
morphs.no <- morphs[-out,]
str(morphs.no)

```

```{r}
#Construct PCA
fit.pca <- princomp(morphs.no[,2:6], cor=TRUE) 
fit.pca
```

```{r}
#Get results of PCA
summary(fit.pca)
plot(fit.pca,type="lines")
loadings(fit.pca)
biplot(fit.pca, col=c("lightgrey","red"))
#The result shows that CP1 and CP2 occupy the most part of variance, so they can get a very accurate view on where it stands in relation to other samples.
```
```{r}
#This is the variable that could measure the overall size.
fit.pca$scores[,1]
```

# 4b
Using ANOVA to test whether the overall size of three morphs are equal. 
```{r}
# Univariate analysis of PC scores
morphs.no$PC1 <- fit.pca$scores[,1]
morphs.no
```

```{r}
#construct model
fit.pc1 <- lm(PC1~morph, data=morphs.no)
#Check assumptions
dwtest(fit.pc1)
#dwtest, p-value = 0.03804, which means residuals autocorrelation.

par(mfrow=c(2,2))
plot(fit.pc1)
#plot is good.
par(mfrow=c(1,1))
#Homogeneity of variances and normality of residuals
leveneTest(fit.pc1)
#The result not significant, which means fullfil assumption

#Test 
anova(fit.pc1)

#Result of anova
summary(fit.pc1)
#P-value is 0.2483, not significant.
```
Through the anova test, the result is not significant, which means the overall size of three morphs are similar (not different).
I don't think the result could indicate three types of morphs eat different types of prey.
#4d
How could the design of this study be improved?
I think we should find an appropriate physical measurement to represent the overall size of species.  
